---
title: "10-read-in"
output: github_document
---


## Import libraries.
```{r}
library(readxl)
library(tidyverse)
library(janitor)
library(ggplot2)
library(tidytext)
library(wordcloud2)
library(RColorBrewer)
```

Library introduction (The parts used in this pipeline):

- **readxl**: read in data in Excel file.
- **tidyverse**: help manage the data frame in tidy sentences.
- **janitor**: clean the column names.
- **ggplot2**: create plots.
- **tidytext**: import stop words data for text analysis and get word sentiment dictionary.
- **wordcloud2**: create word cloud graph.
- **RColorBrewer**: set color palette for graphes.

## Read in raw data and make the column names.

```{r}
for(i in seq(2000,2009,1)){
  df <- read_excel("Analysis Exercise.xls", sheet = as.character(i))
  names(df) <- df[1,]
  df <- df[-1,] %>% 
    clean_names() %>% 
    mutate(year = i)
  assign(paste0("df_", i), df)
}
```

In this Excel file, the data of different years was divided to sub-tables, therefore, I need to read in them one by one. 

And the column names in each table were not in the first row but second row. So I removed the fist row and transferred the second row to be the column names. Then, I used `clean_names()` function to make the names lowercase without blank space.

The function `assign` can help me use different names to reserve the data.

After that, I merged all of the data.
```{r}
df <- 
  bind_rows(df_2000, df_2001, df_2002, df_2003, df_2004, df_2005, df_2006, df_2007, df_2008, df_2009)
```

Import the stop words for text analysis in the EDA part.
```{r}
data("stop_words")
```

