---
title: "R Notebook for md"
output: github_document
---

# 10-read-in

## Import libraries.
```{r}
library(readxl)
library(tidyverse)
library(janitor)
library(ggplot2)
library(tidytext)
library(wordcloud2)
library(RColorBrewer)
```

Library introduction (The parts used in this pipeline):

- **readxl**: read in data in Excel file.
- **tidyverse**: help manage the data frame in tidy sentences.
- **janitor**: clean the column names.
- **ggplot2**: create plots.
- **tidytext**: import stop words data for text analysis and get word sentiment dictionary.
- **wordcloud2**: create word cloud graph.
- **RColorBrewer**: set color palette for graphes.

## Read in raw data and make the column names.

```{r}
for(i in seq(2000,2009,1)){
  df <- read_excel("Analysis Exercise.xls", sheet = as.character(i))
  names(df) <- df[1,]
  df <- df[-1,] %>% 
    clean_names() %>% 
    mutate(year = i)
  assign(paste0("df_", i), df)
}
```

In this Excel file, the data of different years was divided to sub-tables, therefore, I need to read in them one by one. 

And the column names in each table were not in the first row but second row. So I removed the fist row and transferred the second row to be the column names. Then, I used `clean_names()` function to make the names lowercase without blank space.

The function `assign` can help me use different names to reserve the data.

After that, I merged all of the data.
```{r}
df <- 
  bind_rows(df_2000, df_2001, df_2002, df_2003, df_2004, df_2005, df_2006, df_2007, df_2008, df_2009)
```

Import the stop words for text analysis in the EDA part.
```{r}
data("stop_words")
```


# 20-feature-engineering

This data set is clean enouph and I don't need to clean the data as normal.

Basic feature engineering on raw merged data:
```{r}
df <- df %>% 
  mutate(weeks_at_number_1 = as.integer(weeks_at_number_1),
         weeks_top_10 = as.integer(weeks_top_10),
         weeks_top_25 = as.integer(weeks_top_25),
         weeks_top_50 = as.integer(weeks_top_50),
         peak_position = as.integer(peak_position),
         bonus_weeks = as.integer(bonus_weeks),
         sub_points = as.integer(sub_points),
         peak_points = as.integer(peak_points),
         total_points = as.integer(total_points),
         peak_year = as.integer(peak_year),
         debut_date = as.integer(debut_date),
         debut_date = as.Date(debut_date, origin="1899-12-30"), 
         peak_date = as.integer(peak_date),
         peak_date = as.Date(peak_date, origin="1899-12-30"),
         weeks_to_reach_peak = as.integer(weeks_to_reach_peak),
         year=as.factor(year)
         )
```

In this chunk, I only need to make all of the column have right data type.

**Hint:** In the date part, because Excel save date by count the distance between the date and 1900-01-01, the date columns in the raw data are the days. I need to tranfer them to be the real date. The `origin` attribute shoule be 1900-01-01, but due to 1900-leap-year-bug of Excel, use 1899-12-30 can get right answer.



# 30-exploratory-data-analysis



In this part, I will answer the following questions to finish the exploratory data analysis(EDA) part.

## What percent of songs that chart never make the Top 10?

For all:
```{r}
never_top10 <- df %>% 
  filter(weeks_top_10 == 0)
nrow(never_top10)/nrow(df)
```

Time series:
```{r}
never_top10_year <- df %>% 
  filter(weeks_top_10 == 0) %>% 
  group_by(year) %>% 
  summarise(never_top10_number = n())

all_year <- df %>% 
  group_by(year) %>% 
  summarise(whole_number = n())

never_top10_year["whole_number"] <- all_year["whole_number"]
never_top10_year %>% 
  mutate(never_top10_ratio = never_top10_number/whole_number) %>% 
  ggplot() +
  geom_col(aes(x=year, y=never_top10_ratio),fill="#125FA0", width = 0.7)+
  theme_classic()+
  labs(x="\nYear", y="Ratio of Never-Top10\n")
```

## What were the top 10 songs of the decade 2000-2009? 

```{r}
df %>% 
  arrange(desc(total_points)) %>% 
  select(song_title, artist, year, total_points) %>%
  head(10)
```


## Who was the top artist of the decade 2000-2009?

(Who were the top 10 artists of the decade 2000-2009?)

```{r}
df %>% 
  group_by(artist) %>% 
  summarise(sum_total_points = sum(total_points, na.rm = TRUE)) %>% 
  arrange(desc(sum_total_points)) %>% 
  head(10)
```

## Based on the chart data, what artist(s) would you call "The Beatles" of the decade 2000-2009?

Actually, this question is really difficult to answer, because "The Beatles" has various of meaning according to different interpretation. 

However, we can try to give some directions by analysing the number of songs on chart and the mean point of the songs on chart of each artist.

```{r}
df %>% 
  group_by(artist) %>%
  summarise(song_number = n()) %>% 
  arrange(desc(song_number)) %>% 
  head(10)
```

```{r}
df %>% 
  group_by(artist) %>%
  summarise(mean_total_points = mean(total_points)) %>% 
  arrange(desc(mean_total_points)) %>% 
  head(10)
```


## What was the most commonly used word in a song title for the decade 2000-2009?

For this question, due to there are many special symbols in the song title, I need to clean the special patterns in the title.
```{r}
title_df <- df["song_title"]
title_df <- title_df %>% mutate(song_title = str_to_lower(song_title))
title_df$song_title <- gsub(pattern = "\\(", replacement =" ", title_df$song_title)
title_df$song_title <- gsub(pattern = "\\)", replacement =" ", title_df$song_title)
title_df$song_title <- gsub(pattern = ",", replacement =" ", title_df$song_title)
title_df$song_title <- gsub(pattern = "n't", replacement =" not ", title_df$song_title)
title_df$song_title <- gsub(pattern = "'s", replacement =" ", title_df$song_title)
title_df$song_title <- gsub(pattern = "'re", replacement =" are ", title_df$song_title)
title_df$song_title <- gsub(pattern = "'m", replacement =" am ", title_df$song_title)
title_df$song_title <- gsub(pattern = "\\\"", replacement =" ", title_df$song_title)
title_df$song_title <- gsub(pattern = "\\.", replacement =" ", title_df$song_title)
title_df$song_title <- gsub(pattern = "-", replacement =" ", title_df$song_title)
title_df$song_title <- gsub(pattern = "_", replacement =" ", title_df$song_title)
title_df$song_title <- gsub(pattern = "!", replacement =" ", title_df$song_title)
title_df$song_title <- gsub(pattern = "\\?", replacement =" ", title_df$song_title)
title_df$song_title <- gsub(pattern = "…", replacement =" ", title_df$song_title)
title_df$song_title <- gsub(pattern = "[0-9]", replacement =" ", title_df$song_title)
title_df$song_title <- gsub(pattern = "/", replacement =" ", title_df$song_title)
title_df$song_title <- gsub(pattern = "&", replacement =" ", title_df$song_title)
title_df$song_title <- gsub(pattern = "#", replacement =" ", title_df$song_title)
title_df$song_title <- gsub(pattern = " +", replacement =" ", title_df$song_title)
```

Then make the data frame become list, which can help the following steps.
```{r}
title_list <- title_df[["song_title"]]
```

Split all of title into scattered words.
```{r}
title_words <- data.frame()
count <- 1
for (i in title_list) {
  title_split <- str_split(string = i, pattern = " ")
  for(j in title_split[[1]]){
    title_words[count,1] <- j
    count <- count + 1
  }
}
names(title_words) <- "word"
```

clean the different forms of the same word.
```{r}
title_words <- title_words %>% 
  mutate(word = case_when(word == "girls" ~ "girl",
                          word == "angels" ~ "angel",
                          word == "days" ~ "day",
                          word == "lovin'" ~ "love",
                          word == "loving" ~ "love",
                          word == "loved'" ~ "love",
                          word == "luv" ~ "love",
                          word == "waiting" ~ "wait",
                          word == "does" ~ "do",
                          word == "lights" ~ "light",
                          word == "things" ~ "thing",
                          word == "comes" ~ "come",
                          word == "dancing" ~ "dance",
                          word == "dancin'" ~ "dance",
                          word == "boys" ~ "boy",
                          word == "burning" ~ "burn",
                          word == "burnin'" ~ "burn",
                          word == "calle" ~ "call",
                          word == "called" ~ "call",
                          word == "calling" ~ "call",
                          word == "call" ~ "call",
                          TRUE ~ word)) 
```

Count the number of each word.
```{r}
word_count <- title_words %>% 
  filter(word != "") %>% 
  anti_join(stop_words,by = "word") %>% 
  group_by(word) %>% 
  summarise(number = n()) %>% 
  arrange(desc(number)) %>% 
  mutate(word = str_to_upper(word))
word_count 
```

Generate the word cloud by `wordcloud2` function.
```{r}
# wordcloud2(head(word_count,100),
#            size = 1, 
#            fontFamily = "Montserrat",
#            fontWeight = "Bold",
#            minRotation = -pi/6, 
#            maxRotation = -pi/6,
#            rotateRatio = 1,
#            color = "#125FA0")
```

## Sentiment analysis of titles

In order to analyse the sentiment of the words in the titles, I got the stop words(meaningless words) dictionary and sentiment dictionary from the `tidytext` library.
```{r}
sentiments <- get_sentiments("nrc")

df_sentiments <- title_words %>% 
  filter(word != "") %>% 
  anti_join(stop_words,by = "word") %>% 
  left_join(sentiments)

df_sentiments_filtered <- df_sentiments %>% 
  filter(!is.na(sentiment)) %>% 
  group_by(sentiment) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```
After removed the stop word, I counted the number of each words in the titles.


```{r}
df_sentiments_filtered %>% 
  ggplot(aes(x = reorder(sentiment, n, function(n) -n), y = n)) + 
  geom_bar(stat = "identity",fill="#125FA0", width = 0.7) + 
  theme_classic() +
  labs(x = "\nSentiments", y="number\n")
```

And we can also count what is the percentage of positive words in the words that have sentiments.
```{r}
words_positive <- df_sentiments %>% 
  filter(sentiment == "positive") %>% 
  group_by(word) %>% 
  summarise(n=n())

words_all_sentiment <- df_sentiments %>% 
  filter(!is.na(sentiment)) %>% 
  group_by(word) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))

nrow(words_positive)/nrow(words_all_sentiment)
```


## Who spent the most weeks at #1 for the decade 2000-2009?

```{r}
df %>% 
  group_by(artist) %>% 
  summarise(top1_time_sum = sum(weeks_at_number_1)) %>% 
  arrange(desc(top1_time_sum)) %>% 
  head(1) 
```

And there are the songs belong to this artist.
```{r}
df %>% 
  filter(artist == "JACKSON, JANET",
         weeks_at_number_1 != 0) %>% 
  arrange(desc(weeks_at_number_1)) %>% 
  select(song_title, artist, year, weeks_at_number_1)
```


## What song spent the most weeks at #1 for the decade 2000-2009?

```{r}
df %>% 
  arrange(desc(weeks_at_number_1)) %>% 
  select(song_title, artist, year, weeks_at_number_1) %>% 
  head(1)
```

## What song peaked at #1 the quickest in the decade 2000-2009?

```{r}
df %>% 
  filter(weeks_at_number_1 != 0) %>% 
  arrange(weeks_to_reach_peak) %>% 
  select(song_title, artist, year, weeks_at_number_1, weeks_to_reach_peak, debut_date, peak_date) %>% 
  head(1)
```

## What song took the longest to reach #1 in the decade 2000-2009?

```{r}
df %>% 
  filter(weeks_at_number_1 != 0) %>% 
  arrange(desc(weeks_to_reach_peak)) %>% 
  select(song_title, artist, year, weeks_at_number_1, weeks_to_reach_peak, debut_date, peak_date) %>% 
  head(1)
```

## Solo men, solo women, groups for the decade 2000-2009 - design a graphic that shows the distribution of songs hitting #1

The solo artist have their full name in the data by the format as "LastName, FirstName", as a result, we can tell whether the data is a single person name by the presence of a comma.
```{r}
df %>% 
  mutate(solo_group = case_when(str_detect(string = artist, pattern = ",") ~ "solo",
                                TRUE ~ "group")) %>% 
  mutate(solo_group = as.factor(solo_group)) %>% 
  filter(weeks_at_number_1 != 0) %>% 
  group_by(solo_group) %>%
  summarise(number = n()) 
```

Actually, there is no built in pie chart function in `ggplot2`. But we can use bar chart function then twisting the y-axis to the polar coordinate system to create a pie chart.

```{r}
df %>% 
  mutate(solo_group = case_when(str_detect(string = artist, pattern = ",") ~ "solo",
                                TRUE ~ "group")) %>% 
  mutate(solo_group = as.factor(solo_group)) %>% 
  filter(weeks_at_number_1 != 0) %>% 
  ggplot()+
  geom_bar(aes(x=1, fill = solo_group))+
  coord_polar(theta = "y")+
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        legend.title = element_blank(),
        panel.background = element_blank()
        )+
  scale_fill_brewer(palette = "Paired")
```


## Solo men, solo women, groups for the decade 2000-2009 - design a graphic that shows the distribution of songs hitting the Top 10

```{r}
df %>% 
  mutate(solo_group = case_when(str_detect(string = artist, pattern = ",") ~ "solo",
                                TRUE ~ "group")) %>% 
  mutate(solo_group = as.factor(solo_group)) %>% 
  filter(weeks_top_10 != 0) %>% 
  group_by(solo_group) %>%
  summarise(number = n())
```

```{r}
df %>% 
  mutate(solo_group = case_when(str_detect(string = artist, pattern = ",") ~ "solo",
                                TRUE ~ "group")) %>% 
  mutate(solo_group = as.factor(solo_group)) %>% 
  filter(weeks_top_10 != 0) %>% 
  ggplot()+
  geom_bar(aes(x=1, fill = solo_group))+
  coord_polar(theta = "y")+
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        legend.title = element_blank(),
        panel.background = element_blank()
        )+
  scale_fill_brewer(palette = "Paired")
```

## What song spent the most time on the charts in the decade 2000-2009)

```{r}
df %>% 
  arrange(desc(weeks_top_50)) %>% 
  select(song_title, artist, year, weeks_top_50) %>% 
  head(3)
```
Actually, TIME OF YOUR LIFE (GOOD RIDDANCE)	never be top 1. A sad story.

## Is there a correlation between weeks spent on the chart and weeks at #1?

```{r}
cor(x = df[["weeks_at_number_1"]],
    y = df[["weeks_top_50"]])
```

```{r}
df %>% 
  ggplot(aes(x=weeks_at_number_1, y=weeks_top_50))+
  geom_jitter(alpha = 0.3, color = "dodgerblue4")+
  geom_smooth(color = "dodgerblue4", se = FALSE)+
  theme_classic()+
  labs(x="\nWeeks at Top 1", y="Weeks at Top 50\n")
```



